---
title: 'RR_Assignment: Course Project 1'
output: html_document
---

1. Code for reading in the dataset and/or processing the data
```{r Load and process data, echo=TRUE}

data <- read.csv("activity.csv", header=T, sep = ",")

#Get general idea of the data.
head(data)
tail(data)
dim(data)
names(data)
str(data)

#Total number of days in data = 61 days
unique(data$date)

# Keep a list of all possible intervals for plotting
uniqueIntervals <- unique(data$interval)

#There are 288 intervals per day
tail(head(data, 289))

#Loop through each interval for each day
sumperday <- c()
for (i in 1:61) {
        startofday <- ((i*288)-287)
        endofday <- (i*288)
        temp <- data[startofday:endofday, 1]
        #add steps taken each day, ignore NA.
        sumperday <- c(sumperday, sum(temp, na.rm = T))
}
head(sumperday)
str(sumperday)
```


2. Histogram of the total number of steps taken each day
```{r Histogram of total steps, echo=TRUE}

hist(sumperday, breaks=20, col="blue", xlab = "Number of Steps", main = "Total Number of Steps per Day")

```

3. Mean and median number of steps taken each day
```{r Mean and median, echo=TRUE}

mean(sumperday)
median(sumperday)

```

4. Time series plot of the average number of steps taken
```{r Time series plot, echo=TRUE}

#Plot average number of steps
avedaily <- tapply(data$steps, data$interval, mean, na.rm=T)
plot(avedaily ~ unique(data$interval), type="l", xlab="5 Min Interval", main="Average number Steps")
head(avedaily)
tail(avedaily)

```

5. The 5-minute interval that, on average, contains the maximum number of steps
```{r Interval with max steps, echo=TRUE}

#Max occurs at 5 min interval segment 8:35 am.
which.max(avedaily)

#Plot average number of steps
avedaily <- tapply(data$steps, data$interval, mean, na.rm=T)
plot(avedaily ~ unique(data$interval), type="l", xlab="5 Min Interval", main="Average number Steps with Max in Red at 835 or 8:35am")
abline(v=850, col="red", lty=3)

```

6. Code to describe and show a strategy for imputing missing data
```{r Strategy for missing data, echo=TRUE}
#There are 2304 rows containing NA, or missing data.
sum(is.na(data$steps))

#Create another copy of data
data2 <- data
totalsteps <- c()

# Loop through each day and interval and fill with mean. 
for (i in 1:61) {
        startofday <- ((i*288)-287)
        endofday <- (i*288)
        temp <- data2[startofday:endofday, 1]
        #add steps taken each day. Remove NA.
        totalsteps <- c(totalsteps, mean(temp, na.rm=T))
}

head(totalsteps)


```

7. Histogram of the total number of steps taken each day after missing values are imputed
```{r Histogram with filled NA, echo=TRUE}

#Histogram of the total number of steps taken each day after missing values are imputed
hist(totalsteps, xlab="Steps", main = "With NA's Filled in", breaks = 20, col = "red")
hist(sumperday, xlab = "Steps", main = "Total Steps, NA's not Filled", breaks = 20, col = "blue")

```

8. Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
```{r Plot weekend weekday, echo=TRUE}

#Make another copy of data to work with.
datNew <- data
dates <- strptime(datNew$date, "%Y-%m-%d")
datNew$date <- dates

# Keep a list of all possible intervals
uniqueIntervals <- unique(datNew$interval)

#Use POSIXlt date function to group days from 0:6 (Sunday thru Saturday)
wdays <- dates$wday

whichDays <- rep(0, 17568)

#Weekdays are 1:5, weekends are 6,0 (Saturday, Sunday)
whichDays[wdays > 0 & wdays < 6] <- 1
whichDays[wdays == 6 | wdays == 0] <- 2
head(whichDays)

#Since whichDays is just a string of 1's and 2's, create a factor with 2 levels.
daysFactor <- factor(whichDays, levels=c(1,2))
datNew$weekWeekend <- daysFactor

#Sort into Weekdays and weekends.
weekdays <- datNew[datNew$weekWeekend == "1", ]
weekends <- datNew[datNew$weekWeekend == "2", ]

#Split into a list for plotting later.
splitWeekdays <- split(weekdays$steps, weekdays$interval)
splitWeekends <- split(weekends$steps, weekends$interval)

#Take mean, but ignore NA's.
meanWeekdays <- sapply(splitWeekdays, mean, na.rm=TRUE)
meanWeekends <- sapply(splitWeekends, mean, na.rm=TRUE)

#Plot Weekday and Weekend data
par(mfcol=c(2,1))
plot(uniqueIntervals, meanWeekdays, type="l",
main="Average number steps per interval - weekdays", 
xlab="Interval", ylab="Ave # steps - weekdays", 
lwd=2, col="blue")
plot(uniqueIntervals, meanWeekends, type="l",
main="Average number steps per interval - weekends", 
xlab="Interval", ylab="Average # steps - weekends", 
lwd=2, col="blue")

```

